M\"obius Inversion: Construction and Applications

### 2017-04-06 Drafts/Draft.tex
Enriched categories etc

\documentclass[12pt]{pom_thesis}
\input{commands_packages}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}
\usepackage{mathrsfs}
\usepackage[euler-digits]{eulervm} %TODO: Maybe change this
\author{Leo Selker}
\advisor{Vin de Silva}
\title{M\"obius Inversion: Construction and Applications}
\renewcommand{\theequation}{\addtocounter{thm}{0}\thethm}
\newcommand{\fix}{\text{fix }}
\newcommand{\cl}{\text{cl }}
\DeclareMathOperator{\cat}{cat}
\DeclareMathOperator{\ps}{\mathscr{P}}
\DeclareMathOperator{\sub}{Sub}
\DeclareMathOperator{\stab}{Stab}
\DeclareMathOperator{\orb}{Orb}
\DeclareMathOperator{\fixx}{fix}
\DeclareMathOperator{\obj}{ob}
\begin{document}

@@

The goal of M\"obius inversion is to generalize the process of reversing sums over discrete, ordered structures called \emph{posets}. If sums like the above are a discrete analogue of integration (cumulatively summing over a region), then M\"obius inversion is discrete differentiation. In the same way that developing the theory of the derivative is useful for solving differential equations where functions are defined in terms of their integral, M\"obius inversion helps us extract values which are defined in terms of their cumulative sums. In addition to building the necessary machinery for M\"obius inversion, we will consider several applications of it, including a somewhat surprising one in data clustering.
\section{Initial Example and Posets}
M\"obius inversion is a general construct. The classical definition of a M\"obius function is a function $\mu$ on the positive integers which satisfies the following for functions $f,g$ on the positive integers:

\begin{equation}\label{init_mob} 
f(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(m)\mu\left(\frac dn\right)
\end{equation}

@@

\begin{defn}
A \emph{partially ordered set}, often called a \emph{poset}, is a set $X$ together with a binary relation $\leq$, such that, for all $a,b,c \in X$:
\begin{itemize}
\item $a \leq a$ (reflexive)
\item If $a \leq b$ and $b \leq a$ then $a = b$ (antisymmetric)
\item If $a \leq b$ and $b \leq c$ then $a \leq c$ (transitive).
\end{itemize}
\end{defn}
Note that, for completeness, a poset should be labeled with both its elements and its relations - e.g. $(X, \leq)$. But often we just write $X$ when the relation is implied.
\begin{defn}
A poset $X$ is \emph{lower finite} if, for any $a \in x$, the set $\{ c \in X: c \leq a\}$ is finite.
\end{defn}
This condition will be necessary for many of the following facts and theorems, to ensure that certain sums have a finite number of nonzero terms.

@@

Note that, because $X$ is lower finite, this sum has a finite number of terms, so is always finite-valued. Using the ``indicator" function $\zeta$ from above, we can rewrite the definition of $f$ to remove the need for the bound on the sum:
\begin{equation}\label{sum_form_intro}
f(a) = \sum_{b \in X}g(b)\zeta(b,a)
\end{equation}
Each value of $f$, therefore, takes on precisely the structure of a dot product. We will use this structure to leverage linear algebra: 
\begin{itemize}
\item Totally order $X$ such that the partial order is respected;
\item Write $g$ as a row vector which contains the values $g$ takes on elements of $X$, with respect to the total ordering;
\item Write $\zeta$ as a matrix whose $i,j$ entry is equal to $\zeta(a_i, a_j)$, where $a_i, a_j$ are the $i$'th and $j$'th elements of $X$ under the total ordering.
\end{itemize}

@@

If the matrix multiplication is explicitly written out, what we get is identical to \ref{sum_form_intro}. %TODO: Fix this numbering

@@

This means that we can let $\mu=\zeta^{-1}$, which we now know exists, and write down the central relationship of M\"obius inversion, which holds for any lower finite poset $X$ and pair of functions $f,g$ on it:
\[
g = g \zeta \iff g = f \mu.
\]

@@

\section{Definition of M\"obius Inversion}
We are now ready to define the M\"obius function on partially ordered sets.
\begin{defn}\label{def_mob_pos}
Let $X$ be a lower finite poset. Let $\zeta$ be the indicator function defined above. Let $\mu = \zeta^{-1}$. Then $\mu$ is the \emph{M\"obius function} on $X$.
\end{defn}
\begin{rmk}
Recall that $\zeta^{-1}$ is the matrix inverse rather than the functional inverse, and that it exists by \ref{inv_ex}.
\end{rmk}
\begin{rmk}\label{sum_form}
We can unpack the linear algebra to get a definition analagous to \ref{sum_form_intro}: We can say that $\mu$ is the unique function from $X\times X$ to $\Rr$ such that, for all $x,y \in X$,
\[
\sum_{z \in X} \mu(x,z)\zeta(z,y) = \delta(x,z) = \sum_{z \in X} \zeta(x,z)\mu(z,y).
\]
\end{rmk}

@@

We now have two equivalent definitions of the M\"obius function, one using matrices (or linear operators) and vectors, and one directly considering elements and sums. We will continue to move back and forth between these two approaches throughout this investigation. This is because although the two definitions are completely equivalent, they each have advantages. Working with sums and elements helps us see the details of what is going on, and reminds us of the structure of our poset. It is also often the perspective from which motivating problems arise. On the other hand, using matrices and vectors greatly simplifies the equations we have to deal with, makes the manipulations themselves easier, and allows us to keep the bigger picture - the idea of inversion - in mind more easily. 
\section{Integer Lattices}
We now turn to a specific class of partially ordered sets, which will come up several times going forward.
\begin{defn}\label{lattice}
For any positive integer $n$, the \emph{integer lattice of dimension $n$}, which we denote $\Nn^n$, is the set of ordered $n$-tuples of non-negative integers. For any two elements $(a_1,...,a_n), (b_1,...,b_n) \in \Nn^n$, we say that $(a_1,...,a_n)\leq (b_1,...,b_n)$ if $a_i \leq b_i$ for each $i = 1,\cdots,n$.
\end{defn}
\begin{rmk}
$(\Nn^n, \leq)$ is a poset for any positive integer $n$. The axioms are easily checked.
\end{rmk}
We turn to several examples.
\begin{examp}
Consider the $n$-dimensional unit cube in $\Rr^n$, placed such that one corner is at the origin and all vertices have non-negative coordinates. Let $S$ denote the set of vertices of the cube. Then we can think of the vertices of the cube as ordered $n$-tuples of zeros and ones. So $S \subseteq \Nn^{n}$.

@@

In this case, if $\leq$ represents the poset relation from \ref{lattice}, then $m \leq n$ if, for each prime $p$, $m$'s factorization contains no more copies of $p$ than $n$'s factorization. This is the same as saying that $m|n$. So the poset relation becomes divisibility. We will explore this example more in the next sections.
\end{examp}
It turns out that several important posets have this lattice structure. So the goal is to determine the M\"obius function on certain subsets of the integer lattice, and then transfer those results to other partially ordered sets. To do so, we first need a lemma which applies to the M\"obius function on any poset.
\begin{lemma}\label{recurse}
Let $X$ be a lower finite partially ordered set, and let $\zeta, \mu$ be defined as usual. Then for all $a, b \in X$,
\[
\mu(a,b) = 
\begin{cases}
0 & a \nleq b\\
\delta(a,b)-\sum_{a \leq c < b}\mu(a,c) & else.
\end{cases}
\]
\end{lemma}
\begin{rmk}
Note that the piecewise form is actually not necessary, because if $a \nleq b$ then the sum will be empty. So it is correct - albeit less clear- to just write
\[
\mu(a,b) = \delta(a,b)-\sum_{a \leq c < b}\mu(a,c).
\]
\end{rmk}
\begin{proof}
We fix $a,b$, and then rearrange terms starting from \ref{sum_form}:
\begin{align*}
\sum_{c \in X}\mu(a,c) \zeta(c,b) &= \delta(a,b)\\
\sum_{a \leq c < b} \mu(a,c) + \mu(a,b) &= \delta(a,b)\\
\mu(a,b) &= \delta(a,b) - \sum_{a \leq c < b} \mu(a,c) 
\end{align*}
\end{proof}
We are now ready to write down $\mu$ on certain subsets of the integer lattice. In order to ensure that the sums involved are predictable, we need an additional condition on the subsets we consider:

\begin{lemma}\label{mu_lattice}
Let $L$ be the $n$-dimensional integer lattice, with the usual poset relation. Let $X$ be a subset of $L$ such that if $x \in X, y \in L$, and $y \leq x$, then $y \in X$. Let $\zeta, \mu$ be defined as usual on $X$. Let $x=(x_1,...,x_n),y = (y_1,...,y_n)$ be elements of the lattice. Then:
\[
\mu(x,y) =
\begin{cases}
0 & x\nleq y\\
0 & \exists i \in \Nn: y_i-x_i \geq 2\\
(-1)^{\sum_{i=1}^ny_i-x_i} & \text{otherwise}
\end{cases}
\]
\end{lemma}

\begin{proof} %TODO: Prove for general integer lattice
If $x \nleq y$, then by \ref{recurse}, $\mu(x,y) = 0$. This is the first case.

@@

Suppose that we have \ref{mu_lattice} for any $x \leq y$ where $\sum_{i=1}^ny_i-x_i \leq k-1$. Let $x \leq y$ be fixed such that $\sum_{i=1}^ny_i-x_i =k$. If $\exists i \in \Nn: y_i-x_i \geq 2$

If $\frac{n_0}{d_0}$ is not divisible by a square, then $\frac{n_0}{d_0} = p_1p_2\cdots p_{k+1}$, where the $p_i$ are distinct primes. Then there is a direct correspondence:
\begin{center}$\{k : d_0|k, k|n_0, k\neq n_0\} \leftrightarrow$ \{Nonempty subsets of $\{p_1,\cdots,p_{k+1}\}$\},\end{center}
where $k$ is obtained by multiplying the remaining prime factors by $m_0$, and the omitted prime factors can be determined by factoring $k$.

This means that we can index the sum from \ref{recurse} using those subsets:
\[
\mu(m_0, n_0) = -\sum_{P \subsetneq \{1,\cdots,k+1\}} \mu(d_0, m_0\cdot \prod_{i \in P}p_i )
\]
Because $\frac{d_0\cdot \prod_{i \in P}p_i}{d_0} = \prod_{i \in P}p_i$ has $|P| < k + 1$ prime factors, we can apply the inductive hypothesis, which tells us that 
\[
\mu(d_0, md_0\cdot \prod_{i \in P}p_i = (-1)^{|P|}.
\]
So we can write
\begin{align*}
\mu(d_0, n_0) &= -\sum_{P \subsetneq \{1,\cdots,k+1\}} (-1)^{|P|}\\
&= -\sum_{i = 1}^{k}(-1)^i{k+1 \choose i}\\
&= (-1)^{k+1}
\end{align*}
with the final equality following from the alternating-sum binomial identity
\[
\sum_{1=1}^{k+1}(-1)^i{k+1 \choose i} = 0.
\]
Finally, we consider the case where $\frac {d_0}{n_0}$ has a squared factor. So we can write 
\[
\frac {d_0}{n_0} = p^2q
\]
where $p$ is prime, and $q$ is some other integer. Note that we can no longer use the previous argument, because the factors of $\frac {d_0}{n_0} = p^2q$ form a set with multiplicity rather than a simple set, so the sum will no longer have standard binomial coefficients.

Let $r$ divide $d_0q$ and be divisible by $d_0$. So $\mu(d_0, r)$ is a term in the sum from \ref{recurse}. Note also that $rp$ divides $d_0qp$, so $\mu(d_0, rp)$ is also a term in the sum. By the inductive hypothesis, $\mu(d_0, rp) = -\mu(d_0, r)$, so the terms cancel. We can pair terms in this way to eliminate all divisors of $d_0pq$. The only remaining terms are of the form $\mu(d_0, d_0p^2k)$. But because $\frac{d_0p^2k}{d_0}$ is divisible by $p^2$, by the inductive hypothesis, these terms are all zero. So all terms in \ref{recurse} cancel, meaning that $\mu(d_0, n_0) = 0$.
\end{proof}

\section{Integers with Divisibility Ordering} %TODO: Change \Nn to \Nn^+
We are now able to justify the existence of, and find, the function $\mu$ from equation \ref{init_mob}, with little work.

We will consider the set $\Nn$ as a poset ordered by divisibility. For our purposes, $0 \notin \Nn$. Note that this poset is lower finite: For any number $n$, the set $\{q \in \Nn: q|n\} \subseteq \{q \in \Nn: q \leq n\}$, and the latter set contains $n$ elements, which is finite (note that $d|n$ means that $d$ divides $n$ without remainder). This means we are able to use M\"obius inversion.

We go through the usual steps of M\"obius inversion and apply \ref{mob_inv}:
\begin{itemize}
\item Let $\zeta: \Nn \times \Nn \rightarrow \Nn$ be defined as follows.
\[
\zeta(d,n) = \begin{cases}1 & d|n \\ 0 & else \end{cases}
\]
\item Let $\mu = \zeta^{-1}$ 
\end{itemize}
Then \ref{mob_inv} gives us:
\[
g(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(d)\mu(d,n)
\]
This is almost what we want, but not quite. Here $\mu$ is a function in two variables, while in \ref{init_mob} it is a function in a single variable. It turns out that values of $\mu(d,n)$ will only depend on $\frac nd$, which will gives us the desired form. To see that, we start by writing down a recursive form which works for any M\"obius function.

\begin{lemma}\label{recurse}
Let $X$ be a lower finite partially ordered set, and let $\zeta, \mu$ be defined as usual. Then for all $a, b \in X$,
\[
\mu(a,b) = 
\begin{cases}
0 & a \nleq b\\
\delta(a,b)-\sum_{a \leq c < b}\mu(a,c) & else.
\end{cases}
\]
\end{lemma}
\begin{rmk}
Note that the piecewise form is actually not necessary, because if $a \nleq b$ then the sum will be empty. So it is correct - albeit less clear- to just write
\[
\mu(a,b) = \delta(a,b)-\sum_{a \leq c < b}\mu(a,c).
\]
\end{rmk}
\begin{proof}
We fix $a,b$, and then rearrange terms starting from \ref{sum_form}:
\begin{align*}
\sum_{c \in X}\mu(a,c) \zeta(c,b) &= \delta(a,b)\\
\sum_{a \leq c < b} \mu(a,c) + \mu(a,b) &= \delta(a,b)\\
\mu(a,b) &= \delta(a,b) - \sum_{a \leq c < b} \mu(a,c) 
\end{align*}
\end{proof}
We are now ready to write down $\mu$ in the context of the divisibility lattice poset.

\begin{lemma}\label{nfact}
Let $\zeta, \mu$ be defined as usual on $\Nn$ ordered by divisibility. Let $d,n \in \Nn$ be given.  Then:
\[
\mu(d,n) =
\begin{cases}
0 & d\nmid n\\
0 & \exists p \in \Nn: p^2|(\frac nd )\\
(-1)^{\#\{p \text{ prime}: p|\frac nd\}} & else
\end{cases}
\]
\end{lemma}
This lemma gives us what we need - a way of writing values of $\mu$ in terms of a single variable. We will prove it by induction on the number of prime factors of $\frac nd$.
\begin{proof} %TODO: Prove for general integer lattice
If $d \nmid n$, then by \ref{recurse}, $\mu(d,n) = 0$. This is the first case.

If $n = d$, then $\frac nd=1$ has zero prime factors. In that case, again by \ref{recurse}, $\mu(n,d) = 1$. This will be our base case for the second and third branches.

Suppose that if $d|n$, and $\frac nd$ has $k$ prime factors (with multiplicity), then \ref{nfact} holds for $d$ and $n$. Suppose that $d_0|n_0$ and $\frac {n_0}{d_0}$ has $k+1$ prime factors. If $\frac{n_0}{d_0}$ is not divisible by a square, then $\frac{n_0}{d_0} = p_1p_2\cdots p_{k+1}$, where the $p_i$ are distinct primes. Then there is a direct correspondence:
\begin{center}$\{k : d_0|k, k|n_0, k\neq n_0\} \leftrightarrow$ \{Nonempty subsets of $\{p_1,\cdots,p_{k+1}\}$\},\end{center}
where $k$ is obtained by multiplying the remaining prime factors by $m_0$, and the omitted prime factors can be determined by factoring $k$.

>>>>

\documentclass[12pt]{pom_thesis}
\usepackage[T1]{fontenc}
\input{commands_packages}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}
\usepackage{mathrsfs}
\usepackage[euler-digits]{eulervm} %TODO: Maybe change this
\author{Leo Selker}
\advisor{Vin de Silva}
\title{M\"obius Inversion: Construction and Applications}
\newcommand{\fix}{\text{fix }}
\newcommand{\cl}{\text{cl }}
\DeclareMathOperator{\cat}{cat}
\DeclareMathOperator{\ps}{\mathscr{P}}
\DeclareMathOperator{\sub}{Sub}
\DeclareMathOperator{\stab}{Stab}
\DeclareMathOperator{\orb}{Orb}
\DeclareMathOperator{\fixx}{fix}
\DeclareMathOperator{\obj}{ob}
\DeclareMathOperator{\im}{Im}
\newcommand{\catname}[1]{\textbf{\textsc{#1}}}

%\theoremstyle{plain}{
%\newtheorem{thm}{Theorem}[chapter]
%\newtheorem{lemma}[thm]{Lemma}
%\newtheorem{prop}[thm]{Proposition}
%\newtheorem{cor}[thm]{Corollary}
%\newtheorem{conj}[thm]{Conjecture}
%}

%\theoremstyle{definition}{
%\newtheorem{defn}[thm]{Definition}
%\newtheorem{examp}[thm]{Example}
%\newtheorem{rmk}[thm]{Remark}
%\newtheorem{fact}[thm]{Fact}
%}
\begin{document}

@@

The goal of M\"obius inversion is to generalize the process of reversing sums over discrete, ordered structures called \emph{posets}. If sums like the above are a discrete analogue of integration (cumulatively summing over a region), then M\"obius inversion is discrete differentiation. In the same way that developing the theory of the derivative is useful for solving differential equations where functions are defined in terms of their integral, M\"obius inversion helps us extract values which are defined in terms of their cumulative sums. In addition to building the necessary machinery for M\"obius inversion, we will consider several applications of it, including a somewhat surprising one in data clustering.
\section{Initial Example and Posets}
M\"obius inversion is a general construct. The classical definition of a M\"obius function is a function $\mu$ on the positive integers which satisfies the following for functions $f,g$ on the positive integers:
\begin{equation}\label{init_mob} 
f(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(d)\mu\left(\frac dn\right)
\end{equation}

@@

\begin{defn}
A \emph{partially ordered set}, often called a \emph{poset}, is a set $X$ together with a binary relation $\leq$, such that, for all $a,b,c \in X$:
\begin{itemize}
\item $a \leq a$ (reflexive)
\item If $a \leq b$ and $b \leq a$ then $a = b$ (antisymmetric)
\item If $a \leq b$ and $b \leq c$ then $a \leq c$ (transitive).
\end{itemize}
\end{defn}
Note that, for completeness, a poset should be labeled with both its elements and its relations - e.g. $(X, \leq)$. But often we just write $X$ when the relation is implied.
\begin{rmk}
Given $x_1, x_2$ in a poset $X$, we write $[x_1, x_2]$ to denote the set 
\[
\{y \in X: x_1 \leq y \leq x_2\}
\]
Replacing one of the square brackets with a parenthesis means we are omitting that element. So
\begin{align*}
[x_1, x_2) &= \{y \in X: x_1 \leq y < x_2\},\\
(x_1, x_2] &= \{y \in X: x_1 < y \leq x_2\}.
\end{align*}
Note that if $x_1 \nleq x_2, [x_1, x_2] = [x_1, x_2) = (x_1, x_2]= \emptyset.$
\end{rmk}
\begin{defn}
A poset $X$ is \emph{lower finite} if, for any $a \in x$, the set $\{ c \in X: c \leq a\}$ is finite.
\end{defn}
This condition will be necessary for many of the following facts and theorems, to ensure that certain sums have a finite number of nonzero terms. Note that in a lower finite poset, $[x_1, x_2]$ is finite for any $x_1, x_2$.

@@

Note that, because $X$ is lower finite, this sum has a finite number of terms, so is always finite-valued. Using the ``indicator'' function $\zeta$ from above, we can rewrite the definition of $f$ to remove the need for the bound on the sum:
\begin{equation}\label{sum_form_intro}
f(a) = \sum_{b \in X}g(b)\zeta(b,a)
\end{equation}
Each value of $f$, therefore, takes on precisely the structure of a dot product. We will use this structure to leverage linear algebra: 
\begin{itemize}
\item Totally order $X$ such that the partial order is respected;
\item Write $g$ as a row vector which contains the values $g$ takes on elements of $X$, with respect to the total ordering;
\item Write $\zeta$ as a matrix whose $i,j$ entry is equal to $\zeta(a_i, a_j)$, where $a_i, a_j$ are the $i$'th and $j$'th elements of $X$ under the total ordering.
\end{itemize}

@@

If the matrix multiplication is explicitly written out, what we get is identical to \eqref{sum_form_intro}.

@@

This means that we can let $\mu=\zeta^{-1}$, which we now know exists, and write down the central relationship of M\"obius inversion, which holds for any lower finite poset $X$ and pair of functions $f,g$ on it:
\[
f = g \zeta \iff g = f \mu.
\]

@@

\section{Definition of M\"obius Inversion}
We are now ready to define the M\"obius function on partially ordered sets.
\begin{defn}\label{def_mob_pos}
Let $X$ be a lower finite poset. Let $\zeta$ be the indicator function defined above. Let $\mu = \zeta^{-1}$. Then $\mu$ is the \emph{M\"obius function} on $X$.
\end{defn}
\begin{rmk}
Recall that $\zeta^{-1}$ is the matrix inverse rather than the functional inverse, and that it exists by \ref{inv_ex}.
\end{rmk}
\begin{rmk}\label{sum_form}
We can unpack the linear algebra to get a definition analagous to \eqref{sum_form_intro}: We can say that $\mu$ is the unique function from $X\times X$ to $\Rr$ such that, for all $x,y \in X$,
\[
\sum_{z \in X} \mu(x,z)\zeta(z,y) = \delta(x,y) = \sum_{z \in X} \zeta(x,z)\mu(z,y),
\]
or
\[
\sum_{z \leq y} \mu(x,z) = \delta(x,y) = \sum_{z \geq x} \mu(z,y).
\]
\end{rmk}

@@

We now have two equivalent definitions of the M\"obius function, one using matrices (or linear operators) and vectors, and one directly considering elements and sums. We will continue to move back and forth between these two approaches throughout this investigation. This is because although the two definitions are completely equivalent, they each have advantages. Working with sums and elements helps us see the details of what is going on, and reminds us of the structure of our poset. It is also often the perspective from which motivating problems arise. On the other hand, using matrices and vectors greatly simplifies the equations we have to deal with, makes the manipulations themselves easier, and allows us to keep the bigger picture - the idea of inversion - in mind more easily. 
\section{Integer Lattices}
We now turn to a specific class of partially ordered sets, which will come up several times going forward.
\begin{defn}\label{lattice}
For any positive integer $n$, the \emph{integer lattice of dimension $n$}, which we denote $\Nn^n$, is the set of ordered $n$-tuples of non-negative integers. For any two elements $(a_1,\dots,a_n), (b_1,\dots,b_n) \in \Nn^n$, we say that $(a_1,\dots,a_n)\leq (b_1,\dots,b_n)$ if $a_i \leq b_i$ for each $i = 1,\cdots,n$.
\end{defn}
\begin{lemma}
$(\Nn^n, \leq)$ is a lower finite poset for any positive integer $n$. 
\end{lemma}
\begin{proof}
The poset axioms are easily checked. We also observe that for any $x \in (\Nn^n, \leq)$,
\[
\#\{y \in (\Nn^n, \leq):y \leq x\} = \prod_{i =1}^n x_i.
\]

This is a finite product of integers, so is finite.
\end{proof}

We turn to several examples.
\begin{examp}
Consider the $n$-dimensional unit cube in $\Rr^n$, placed such that one corner is at the origin and all vertices have non-negative coordinates. Let $S$ denote the set of vertices of the cube. Then we can think of the vertices of the cube as ordered $n$-tuples of zeros and ones. So $S \subseteq \Nn^{n}$.

@@

In this case, if $\leq$ represents the poset relation from \ref{lattice}, then $m \leq n$ if, for each prime $p$, $m$'s factorization contains no more copies of $p$ than $n$'s factorization. This is the same as saying that $m|n$. So the poset relation becomes divisibility. We will explore this example more in the next sections.
\end{examp}
It turns out that several important posets have this lattice structure. So the goal is to determine the M\"obius function on certain subsets of the integer lattice, and then transfer those results to other partially ordered sets. To do so, we first need a lemma which applies to the M\"obius function on any poset.
\begin{lemma}\label{recurse}
Let $X$ be a lower finite partially ordered set, and let $\zeta, \mu$ be defined as usual. Then for all $a, b \in X$,
\[
\mu(a,b) = 
\begin{cases}
0 & a \nleq b\\
\delta(a,b)-\sum_{c \in [a,b)}\mu(a,c) & else.
\end{cases}
\]
\end{lemma}
\begin{rmk}
Note that the piecewise form is actually not necessary, because if $a \nleq b$ then the sum will be empty. So it is correct - albeit less clear- to just write
\[
\mu(a,b) = \delta(a,b)-\sum_{c \in [a,b)}\mu(a,c).
\]
\end{rmk}
\begin{proof}
We fix $a,b$, and then rearrange terms starting from \ref{sum_form}:
\begin{align*}
\sum_{c \in X}\mu(a,c) \zeta(c,b) &= \delta(a,b)\\
\sum_{c \in [a,b)} \mu(a,c) + \mu(a,b) &= \delta(a,b)\\
\mu(a,b) &= \delta(a,b) - \sum_{c \in [a,b)} \mu(a,c) 
\end{align*}
\end{proof}
We are now ready to write down $\mu$ on certain subsets of the integer lattice. In order to ensure that the sums involved are predictable, we need an additional condition on the subsets we consider:

\begin{lemma}\label{int_lattice}
Let $L$ be the $n$-dimensional integer lattice, with the usual poset relation. Let $\zeta, \mu$ be defined as usual on $L$. Let $x=(x_1,\dots,x_n),y = (y_1,\dots,y_n)$ be elements of $L$. Then:
\[
\mu(x,y) =
\begin{cases}
0 & x\nleq y\\
0 & \exists i \in \Nn: y_i-x_i \geq 2\\
(-1)^{\sum_{i=1}^ny_i-x_i} & \text{otherwise}
\end{cases}
\]
\end{lemma}

\begin{proof}
If $x \nleq y$, then by \ref{recurse}, $\mu(x,y) = 0$. This is the first case.

@@

Suppose that we have \ref{int_lattice} for any $x \leq y$ where $\sum_{i=1}^ny_i-x_i \leq k-1$. Let $x \leq y$ be fixed such that $\sum_{i=1}^ny_i-x_i =k$. Suppose that for some $j \in \Nn, y_j-x_j \geq 2$. Let $y' = (y_1,\dots,y_j-1,\dots,y_n)$. Note that $x \leq y' < y$. So we can decompose the sum from $\ref{int_lattice}$. Let
\begin{align*}
\alpha &= \sum_{x \leq z \leq y'} \mu(x,z)\\
\beta &= \sum_{y' < z < y}\mu(x,z).\\
\end{align*}
Then
\begin{align*}
\mu(x,y) &= -\sum_{x \leq z \leq y}\mu(x,z)\\
&= -\sum_{x \leq z \leq y'} \mu(x,z) - \sum_{y' < z < y}\mu(x,z)\\
&= -(\alpha + \beta).
\end{align*}
We note that
\[\alpha = \sum_{z \in L} \mu(x,z)\zeta(z,y').
\]
So by \ref{sum_form}, $\alpha=0$.

Recall that for all $i \neq j, y'_i = y_i$. This means that for any $z \in L$ where $z>y'$ and $y > z, z_j = y_j$. So $z_j-x_j \geq 2$. By the inductive hypothesis, this implies that $\mu(z,x)=0$. So all the terms of $\beta$ are zero, so $\beta=0$. This implies that \[\mu(z,y)=-(\alpha+\beta)=0.\]

Finally, suppose that for all $i = 1,\dots,n, y_i-x_i \leq 1$. Because $\sum_{i=1}^ny_i-x_i =k$, there must be $k$ indices where $x$ and $y$ differ by one, while at the other $n-k$ indices they are equal. So for each $j = 0,\dots,k-1$, there are $k \choose j$ elements $z$ of $X$ where $x \leq z < y$ and $\sum_{i = 1}^n y_i-z_i = j$. For each of these, by the inductive hypothesis, $\mu(x,z) = (-1)^j$. So, using \ref{sum_form}, we can write
\begin{align*}
\mu(x,y)& = -\sum_{j = 1}^k {k \choose j} (-1)^j\\
&= (-1)^k.
\end{align*}
\end{proof}
This result will be powerful for determining values of $\mu$ on various posets whose structure is similar to an integer lattice. 
\begin{thm}\label{mapping}
Let $(X, \leq_X),(Y, \leq_Y)$ be lower finite posets. Let $x_1 \leq x_2 \in S$ be given. Suppose that there exists a bijective map
\[
\phi: [x_1, x_2] \leftrightarrow [\phi(x_1), \phi(x_2)]
\]
such that for all $z_1, z_2 \in [x_1, x_2]$,
\[ 
z_1 \leq_X z_2 \iff \phi(z_1) \leq_Y \phi(z_2).
\] 
Let $\mu_X, \mu_Y$ denote the M\"obius functions on $X$ and $Y$ respectively. Then:
\[
\mu_X(x_1, x_2) = \mu_Y(\phi(x_1), \phi(x_2)).
\]
\end{thm}
This theorem tells us that if we can find a portion of a familiar poset which has identical structure to a portion of an unknown poset, we can transfer what we know about values of $\mu$ on that portion from one to the other.
\begin{proof}
We induct on $\#[x_1, x_2]$. Because $x_1 \leq x_2$, if $[x_1, x_2] = 1$, then $x_1 = x_2$. So by \ref{recurse}, $\mu_X(x_1, x_2) = 1$. Because $\phi(x_1) = \phi(x_2), \mu_Y(\phi(x_1), \phi(x)2) = 1$ also.

Suppose the theorem is true if $\#[x_1, x_2] \leq k-1$ for some integer $k$. Fix $x_1, x_2$ such that $\#[x_1, x_2] = k$. Then by \ref{recurse},
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{y \in [\phi(x_1), \phi(x_2)]}\mu_Y(\phi(x_1), y)
\]
Because $\phi$ is an isomorphism which preserves $\leq$, we can rewrite this sum:
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{x \in [x_1, x_2]}\mu_Y(\phi(x_1), \phi(x))
\]
We can restrict $\phi$ to $[\phi(x_1), \phi(x)]$ to satisfy the hypotheses for all intermediate elements $x$. So we can apply the inductive hypothesis to all the terms of this sum:
\begin{align*}
\mu_Y(\phi(x_1), \phi(x_2)) &= -\sum_{x \in [x_1, x_2]}\mu_X(x_1, x)\\
&= \mu_X(x_1, x_2)
\end{align*}
\end{proof}
The following corollary follows immediately if we let $\phi$ be the inclusion map:
\begin{cor}\label{subset}
Let $X$ be a poset. Let $I$ be an interval in $X$. In other words, let $I = [z_1, z_2]$ for some $z_1, z_2 \in X$. Then
\[
\mu_I = \mu_X|_I.
\]
\end{cor}
In the next section, we explore a poset where this fact will be extremely useful.

\section{Integers with Divisibility Ordering}
We are now able to justify the existence of, and find, the function $\mu$ from \eqref{init_mob}, with little work.

We will consider the poset $(\Zz^{+}, |)$. This is the set of positive integers ordered by divisibility (recall that $d|n$ if $d$ divides $n$ without remainder). Note that this poset is lower finite: For any number $n$, the set $\{q \in \Zz^{+}: q|n\} \subseteq \{q \in \Zz^{+}: q \leq n\}$, and the latter set contains $n$ elements, which is finite (note that $d|n$ means that $d$ divides $n$ without remainder). This means we are able to use M\"obius inversion. Rather than compute the M\"obius function directly, we can use \ref{int_lattice} and \ref{mapping} to reuse what we already know.

\begin{thm}\label{mu_div}
Let $\zeta, \mu$ be defined as usual on the poset $(\Zz^+, |)$. Then for all $d,n \in \Zz^+$,
\[
\mu(d,n) =
\begin{cases}
0 & \nmid n \\
0 & \exists p: p^2|\frac nd\\
(-1)^{\#\{p|\frac nd : p \text{ prime}\}} & otherwise
\end{cases}
\]
\end{thm}

\begin{proof}
Let $d,n \in \Zz^+$ be given. We set up the isomorphism required for \ref{mapping}: Index the primes $p_1, p_2, \dots$. So $p_1=2, p_2=3$, and so on. Let $k$ be the index of the largest prime which divides either $d$ or $n$. Then let $\phi : [d,n] \rightarrow \Nn^k$ be defined by
\[
 p_1^{x_1}p_2^{x_2}\cdots p_k^{x_k} \mapsto (x_1, \cdots, x_k).
\]
We claim that $\phi$ is a bijection from $[d,n]$ to $\im(\phi)$.

Because prime factorizations are unique, this map is well-defined and injective. Let $m_1, m_2 \in [d,n]$ be given. Then $m_1|m_2$ if and only if the multiplicity of each prime $p_i$ in the factorization of $m_1$ is equal to or smaller than $p_i$'s multiplicity in the factorization of $m_2$. This is equivalent to checking whether or not $\phi(m_1)_i \leq \phi(m_2)_i$, which is the definition of $\leq$ on $\Nn^k$. So this map preserves $\leq$. This implies that $\im(\phi) \subseteq [\phi(d), \phi(n)]$, which allows us to check the final condition, surjectivity.

Let $x = (x_1,...,x_k) \in [\phi(d), \phi(n)]$ be given. Then $\phi(p_1^{x_1}\cdots p_k^{x_k}) = x$. By above, because $\phi$ preserves $\leq$, $p_1^{x_1}\cdots p_k^{x_k} \in [d,n]$. So $x \in \im(\phi)$.

So $\phi:[d,n] \leftrightarrow [\phi(d), \phi(n)]$ is a bijection.

This allows us to use \ref{mapping}, and the result immediately follows.
\end{proof}

This definition of $\mu$ depends only on $\frac nd$, which means that we can abuse notation to write $\mu(\frac nd)$ instead of $\mu(d,n)$. This allows us to use \ref{mob_inv} to arrive at \eqref{init_mob}:
\[
f(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(d)\mu\left(\frac nd\right).
\]
\section{Euler's Totient} %TODO: Fix \Zz^+
Euler's Totient is a famous function in number theory, which counts the how many numbers are relatively prime to a particular value. We denote the function by $\varphi : \Nn \rightarrow \Nn$, and its definition is:
\[
\varphi(n) = \#\{1 \leq m \leq n : (m,n) = 1\},
\]
where $(m,n)$ denotes the greatest common divisor of $m,n$.
